### 2018-04-12 paper/background.tex

\subsubsection{Structure}
% definition of CHC
A Constrained Horn Clause is a logical implication where the
antecedent is called the body and the consequent is called the head.
%
The body is a conjunction of a logical formula, called the fact,
and a vector of uninterpreted predicates. The fact is an arbitrary
formula in some background logic, such as linear integer arithmetic.
%
The uninterpreted predicates are applied to variables which may or may
not appear in the fact.
%
A head can be either an uninterpreted predicate applied to variables or \cc{False}.
%
A clause where the head is \cc{False} is called a query. A CHC can be
defined structurally:
\begin{align*}
\cc{chc} &\Coloneqq \cc{head} \gets \cc{body} \\
\cc{head} &\Coloneqq \cc{False} \\&\mid \cc{pred}  \\
\cc{body} &\Coloneqq \varphi \wedge \cc{preds} \\
\cc{preds} &\Coloneqq \cc{True} \\&\mid \cc{pred} \wedge \cc{preds} \\
\cc{pred} &\Coloneqq \textsf{an uninterpreted predicate applied to variables} \\
\varphi &\Coloneqq \textsf{a formula that all variables are vocabulary} \\
\end{align*}
%
% define a body function, head function, and recursion-free
For a given CHC $C$, $\bodyof{C}$ denotes the vector of uninterpreted
predicates in the body and $\consof{C}$ denotes the fact in the
body.
%
It is possible for an uninterpreted predicate to appear multiple times
in the body of $C$.
%
If $C$ is not a query, then $\headof{C}$ denotes the uninterpreted
predicate in the head.
%

%define a CHC system
A CHC system is a set of CHCs where exactly one clause is a query.
%
For a given CHC system $S$, $\predof{S}$ denotes the set of all
uninterpreted predicates.

% define dependency predicates
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates
  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
  such that $\cc{P} = \headof{C}$ and $\cc{Q} \in \bodyof{C}$, then
  $\cc{Q}$ is a predicate dependency of $\cc{P}$.
\end{defn}
%
\begin{ex}
  In $\mcchc$, because $\cc{L}_4$ is in the body of clause (4) and
  $\cc{L}_8$ is the head of clause (4), $\cc{L}_4$ is a predicate
  dependency of $\cc{L}_8$.
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\depsof{P}$ denotes the set of all predicate dependencies of $\cc{P}$
in $\cc{S}$.
%

%define transitive dependency predicates
\begin{defn}
  Given a CHC system $\cc{S}$ and three uninterpreted predicates
  $\cc{P}, \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in
  \depsof{P}$ then \cc{Q} is a transitive predicate dependency of
  \cc{P}.
  %
  If \cc{Q} is a transitive predicate dependency of \cc{P} and \cc{R}
  is a transitive predicate dependency of \cc{Q}, then $\cc{R}$ is a
  transitive predicate dependency of $\cc{P}$.
\end{defn}
%
\begin{ex}
  In $\mcchc$, because $\cc{L}_4$ is a predicate dependency of
  $\cc{L}_8$, $\cc{L}_4$ is a transitive predicate dependency of
  $\cc{L}_8$.
  %
  And because $\cc{L}_8$ is an transitive predicate dependency of
  $\cc{L}_9$, $\cc{L}_4$ is a transitive predicate dependency of
  $\cc{L}_9$.
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\tdepsof{P}$ denotes the set of all transitive predicate dependencies
of $\cc{P}$ in $\cc{S}$.
%

%define siblings
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates
  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
  such that $\cc{P} \in \bodyof{C}$ and $\cc{Q} \in \bodyof{C}$, then
  $\cc{Q}$ and $\cc{P}$ are siblings.
\end{defn}
%
\begin{ex}
  Because uninterpreted predicates $\cc{L}_9$ and $\cc{db}$ both
  appear in the body of clause (6), $\cc{L}_9$ and $\cc{db}$ are
  siblings.
  %
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\siblingof{P}$ denotes the set of all siblings of $\cc{P}$
in $\cc{S}$.
%
For a given CHC system $\cc{S}$, if there are no uninterpreted
predicate $\cc{P} \in \predof{S}$ such that $\cc{P} \in \tdepsof{P}$,
then $\cc{S}$ is a \emph{recursion-free} CHCs system.
%

@@

% interpolants:
\subsection{Logical interpolation}
\label{sec:itps}
%
All objects defined in this paper are defined over a fixed space of
first-order logical variables $X$.
%
In particular, for a theory $\mathcal{T}$, the space of $\mathcal{T}$
formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
%
For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
$\varphi$) is denoted $\vocab(\varphi)$.
%
% define models, satisfaction, entailment
For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
\emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
\entails \varphi$.

>>>>

\subsubsection{Structure}
% definition of CHC
A Constrained Horn Clause is a logical implication where the
antecedent is called the body and the consequent is called the head.
%
The body is a conjunction of a logical formula, called the fact,
and a vector of uninterpreted predicates. The fact is an arbitrary
formula in some background logic, such as linear integer arithmetic.
%
The uninterpreted predicates are applied to variables which may or may
not appear in the fact.
%
A head can be either an uninterpreted predicate applied to variables or $False$.
%
A clause where the head is $False$ is called a query. A CHC can be
defined structurally:
\begin{align*}
\cc{chc} \Coloneqq&~\cc{head} \gets \cc{body} \\
\cc{head} \Coloneqq&~False \\
  \mid&~\cc{pred} \\
\cc{body} \Coloneqq&~\varphi \wedge \cc{preds} \\
\cc{preds} \Coloneqq&~True \\
  \mid&~\cc{pred} \wedge \cc{preds} \\
\cc{pred} \Coloneqq&~\textsf{\emph{an uninterpreted predicate applied to variables}} \\
\varphi \Coloneqq&~\textsf{\emph{a formula}} \\
\end{align*}
%
% define a body function, head function, and recursion-free
For a given CHC $C$, $\bodyof{C}$ denotes the vector of uninterpreted
predicates in the body and $\consof{C}$ denotes the fact in the
body.
%
If $C$ is not a query, then $\headof{C}$ denotes the uninterpreted
predicate in the head.
%
%define a CHC system
A CHC system is a set of CHCs where exactly one clause is a query.
%
For a given CHC system $S$, $\predof{S}$ denotes the set of all
uninterpreted predicates.

To explain the structure of a CDD system, we need terminology that relates
predicates in a CHC system including the terms \emph{predicate dependency},
\emph{transitive predicate dependency}, and \emph{sibling}.
% define dependency predicates
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates
  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
  such that $\cc{P} = \headof{C}$ and $\cc{Q} \in \bodyof{C}$, then
  $\cc{Q}$ is a predicate dependency of $\cc{P}$.
\end{defn}
%
\begin{ex}
  In $\mcchc$, because $\cc{L}_4$ is in the body of clause (4) and
  $\cc{L}_8$ is the head of clause (4), $\cc{L}_4$ is a predicate
  dependency of $\cc{L}_8$.
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\depsof{P}$ denotes the set of all predicate dependencies of $\cc{P}$
in $\cc{S}$.
%
%define transitive dependency predicates
\begin{defn}
  Given a CHC system $\cc{S}$ and three uninterpreted predicates
  $\cc{P}, \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in
  \depsof{P}$ then \cc{Q} is a transitive predicate dependency of
  \cc{P}.
  %
  If \cc{Q} is a transitive predicate dependency of \cc{P} and \cc{R}
  is a transitive predicate dependency of \cc{Q}, then $\cc{R}$ is a
  transitive predicate dependency of $\cc{P}$.
\end{defn}
%
\begin{ex}
  In $\mcchc$, because $\cc{L}_4$ is a predicate dependency of
  $\cc{L}_8$, $\cc{L}_4$ is a transitive predicate dependency of
  $\cc{L}_8$.
  %
  And because $\cc{L}_8$ is an transitive predicate dependency of
  $\cc{L}_9$, $\cc{L}_4$ is a transitive predicate dependency of
  $\cc{L}_9$.
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\tdepsof{P}$ denotes the set of all transitive predicate dependencies
of $\cc{P}$ in $\cc{S}$.
%
%define siblings
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates
  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
  such that $\cc{P} \in \bodyof{C}$ and $\cc{Q} \in \bodyof{C}$, then
  $\cc{Q}$ and $\cc{P}$ are siblings.
\end{defn}
%
\begin{ex}
  Because uninterpreted predicates $\cc{L}_9$ and $\cc{db}$ both
  appear in the body of clause (7), $\cc{L}_9$ and $\cc{db}$ are
  siblings.
  %
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\siblingof{P}$ denotes the set of all siblings of $\cc{P}$
in $\cc{S}$.
%
For a given CHC system $\cc{S}$, if there are no uninterpreted
predicate $\cc{P} \in \predof{S}$ such that $\cc{P} \in \tdepsof{P}$,
then $\cc{S}$ is a \emph{recursion-free} CHCs system.
%

@@

% interpolants:
\subsection{Logical interpolation}
\label{sec:itps}
%
All logical objects in this paper are defined over a fixed space of first-order
variables, $X$.
%
For a theory $\mathcal{T}$, the space of $\mathcal{T}$
formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
%
For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
$\varphi$) is denoted $\vocab(\varphi)$.
%
% define models, satisfaction, entailment
For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
\emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
\entails \varphi$.

### 2018-04-12 paper/fig/hierarchy.tex

\scalebox{0.75}{%
\begin{tikzpicture}
\begin{scope}
  \node (top) at (0,0) {\begin{tabular}{c}All Systems\\ \textsc{Undecidable}\end{tabular}};
    \node[below = 0.5 of top] (rec) {\begin{tabular}{c}Recursion Free\\
  \textsc{co-Nexptime Complete}\end{tabular}};
  \node[below = 0.5 of rec] (dir) {\begin{tabular}{c}Directly Solvable\\ \textsc{co-NP}\end{tabular}};
  \node[label={0:\{\textsc{Shara}\}}, draw, rounded corners=0.5cm,
      below = 0.5 of dir] (cdd) {\begin{tabular}{c}Clause-Dependent-Disjoint\\ \textsc{co-NP}\end{tabular}};
  \node[label={0:\{\textsc{Ufo}\}},
      below left = 0.5 and 0.5 of cdd] (lin) {\begin{tabular}{c}Linear\\ \textsc{co-NP}\end{tabular}};
  \node[label={180:\{\textsc{Eldarica}\cite{rummer13b}\}},
      below right = 0.5 and 0.5 of cdd] (bd) {\begin{tabular}{c}Body Disjoint\\ \textsc{co-NP}\end{tabular}};
  \node[label={180:\{\textsc{Duality}\cite{bjorner13}\}},
      below = 0.5 of bd] (tree) {\begin{tabular}{c}Tree\\ \textsc{co-NP}\end{tabular}};
\end{scope}
\begin{scope}[every edge/.style={draw=black, very thick}]
  \path[-] (tree) edge (bd);
  \path[-] (bd) edge (cdd);
  \path[-] (lin) edge (cdd);
  \path[-] (cdd) edge (dir);
  \path[-] (dir) edge (rec);
  \path[-] (rec) edge (top);
\end{scope}
\end{tikzpicture}
}

>>>>

\begin{tikzpicture}
\begin{scope}
  \node at (0, 0) (rec) {%
    \begin{tabular}{c}
      Recursion Free\\\textsc{co-Nexptime Complete}
    \end{tabular}
  };

  \node[below = 0.5 of rec] (dir) {%
    \begin{tabular}{c}
      Directly Solvable\\\textsc{co-NP}
    \end{tabular}
  };

  \node[draw, rounded corners=0.5cm, below = 0.5 of dir] (cdd) {%
    \begin{tabular}{c}
      Clause-Dependence Disjoint\\\textsc{co-NP}
    \end{tabular}
  };

  \node[below left = 0.5 and -1.1 of cdd] (lin) {%
    \begin{tabular}{c}
      Linear\\\textsc{co-NP}
    \end{tabular}
  };

  \node[below right = 0.5 and -1.1 of cdd] (bd) {%
    \begin{tabular}{c}
      Body-Disjoint\\\textsc{co-NP}
    \end{tabular}
  };

\end{scope}
\begin{scope}[every edge/.style={draw=black, very thick}]
  \path[-] (bd) edge (cdd);
  \path[-] (lin) edge (cdd);
  \path[-] (cdd) edge (dir);
  \path[-] (dir) edge (rec);
\end{scope}
\end{tikzpicture}

### 2018-04-12 paper/overview.tex
added motivation for definitions

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.3\textwidth] {%
      \caption{\cc{dblAbs}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    {\lstinputlisting[
      basicstyle=\small,
      xleftmargin=\parindent,
      numbers=left,
      morekeywords={def, return, if, else, assert}]{code/dblAbs.java}
        % \input{code/dblAbs.java}}
    }
    %
    \ffigbox[.66\textwidth]{%
      \caption{A CHC system that models the safety condition of $\cc{dblAbs}$,
    named $\mcchc$.}}{%
      \label{fig:chc}
      \begin{align}
        % semantic constraint of dbl:
        \cc{dbl}(\cc{x}, \cc{d}) &\gets \cc{d} = 2 * \cc{x} \\
        % semantic constraint of line 3:
        \cc{L}_4(\cc{n}, \cc{abs}) &\gets  \cc{abs} = 0 \\
        % semantic constraint of line 5 (then branch)
        \cc{L}_6(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
        % semantic constraint of line 6 (else branch)
        \cc{L}_8(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
        % assign branch
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_6(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_8(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
        % semantic constraint of main procedure:
        \cc{main}(\cc{n},\cc{res}) &\gets \cc{L}_9(\cc{n}, \cc{abs'}) \land
        \cc{dbl}(\cc{x}, \cc{d})
        \land \cc{abs'} = \cc{x} \land \cc{res} = \cc{d} \\
        %
        \bot &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
      \end{align}
    }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{dblAbs} is a small procedure that doubles the absolute value of
its input and stores the result in \cc{res}.
% state the safety problem:
The program also asserts that \cc{res} is greater than or equal to $0$
before exiting.
% talk about solving Horn Clauses:
Verifying this assertion reduces to solving a recursion-free
CHC system over a set of uninterpreted predicates that represent
the control locations in \cc{dblAbs}.
%
In particular, one such system, $\mcchc$, is shown in
\autoref{fig:chc}.
%
%
While $\mcchc$ has been presented as the result
of a translation from \cc{dblAbs}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

@@

% tree solver
Body-disjoint (or derivation tree)
systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
a class of recursion-free CHC system where each uninterpreted
predicate appears in the body of at most one clause and appears in
such a clause exactly once.
%
Such systems cannot model a program with multiple control paths that
share a common pre-subpath, typically modeled as a CHC system with a
uninterpreted predicate that occurs in the body of multiple clauses.
%
$\mcchc$ is not a body-disjoint system because $\cc{L}_4$ shows in the
body of both clause (3) and clause (4).
%
In order to find an interpretation for $\mcchc$, a solver which uses
tree-structure systems would have to create a fresh copy of
$\cc{L}_4$, as well copies for all relational predicates that
$\cc{L}_4$ depends on.

>>>>

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.3\textwidth] {%
      \caption{\cc{dblAbs}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    {\lstinputlisting[
      basicstyle=\small,
      xleftmargin=\parindent,
      numbers=left,
      morekeywords={def, return, if, else, assert}]{code/dblAbs.java}
        % \input{code/dblAbs.java}}
    }
    %
    \ffigbox[.66\textwidth]{%
      \caption{A CHC system that models the safety condition of $\cc{dblAbs}$,
    named $\mcchc$.}}{%
      \label{fig:chc}
      \begin{align}
        % semantic constraint of dbl:
        \cc{dbl}(\cc{x}, \cc{d}) &\gets \cc{d} = 2 * \cc{x} \\
        % semantic constraint of line 3:
        \cc{L}_4(\cc{n}, \cc{abs}) &\gets  \cc{abs} = 0 \\
        % semantic constraint of line 5 (then branch)
        \cc{L}_6(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
        % semantic constraint of line 6 (else branch)
        \cc{L}_8(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
        % assign branch
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_6(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_8(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
        % semantic constraint of main procedure:
        \cc{main}(\cc{n},\cc{res}) &\gets \cc{L}_9(\cc{n}, \cc{abs'}) \land
        \cc{dbl}(\cc{x}, \cc{d})
        \land \cc{abs'} = \cc{x} \land \cc{res} = \cc{d} \\
        %
        False &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
      \end{align}
    }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{dblAbs} is a small procedure that doubles the absolute value of
its input and stores the result in \cc{res}.
% state the safety problem:
The program also asserts that \cc{res} is greater than or equal to $0$
before exiting.
% talk about solving Horn Clauses:
Verifying this assertion reduces to solving a recursion-free
CHC system over a set of uninterpreted predicates that represent
the control locations in \cc{dblAbs}.
%
In particular, one such system, $\mcchc$, is shown in
\autoref{fig:chc}.
%
%
While $\mcchc$ has been presented as the result
of a translation from \cc{dblAbs}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

@@

% tree solver
Body-disjoint (or derivation tree)
systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
a class of recursion-free CHC system where each uninterpreted
predicate appears in the body of at most one clause and appears in
such a clause exactly once.
%
Such systems cannot model a program with multiple control paths that
share a common pre-subpath, typically modeled as a CHC system with a
uninterpreted predicate that occurs in the body of multiple clauses.
%
$\mcchc$ is not a body-disjoint system because $\cc{L}_4$ shows in the
body of both clause (3) and clause (4).
%
In order to solve for $\mcchc$, a solver which uses
body-disjoint systems would have to create a fresh copy of
$\cc{L}_4$, as well copies for all clauses and relational predicates that
$\cc{L}_4$ depends on.
